---
title: "Human Activity Recognition - Practical Machine Learning Project"
author: "Daniel Mendalka"
date: "March 21, 2015"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
library(caret)
library(randomForest)
library(doParallel)
```

# Data
The data used in this project: 

* [training](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)
* [testing](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

## Loading and cleaning

```{r}
na.strings <- c('NA', '', '#DIV/0!')
training <- read.csv("pml-training.csv", na.strings=na.strings)
# testing <- read.csv("pml-testing.csv", na.strings=na.strings)
```

Data contains `r dim(training)[1]` rows with `r dim(training)[2]` variables.

From initial data we can remove columns containing:

* data unreleated to performed excercies (`X`, `user_name`, `raw_timestamp_part1`, `raw_timestamp_part2`, `cvtd_timestamp`)
* mostly NA values
* only one unique value

```{r}
    predictors <- names(training)
    
    predictors <- predictors[colSums(is.na(training)) == 0]
    
    useless <- c('X', 'user_name', 'raw_timestamp_part1', 'raw_timestamp_part2', 'cvtd_timestamp')
    predictors <- predictors[!predictors %in% useless]
    
    training <- subset(training, select=predictors)
    training <- training[-nearZeroVar(training)]
```

List of predictors after cleaning data:
```{r}
names(training)
```

# Model

```{r}
```

To train our model we can use random forest:

```{r, echo=FALSE}
if (file.exists('model.RData')) {
    load('model.RData')
} else {
    cl <- makeCluster(detectCores())
    registerDoParallel(cl)
}
```
```{r}
if (!"model" %in% ls()) {
    
    set.seed(1)
    model <- train(classe ~ ., data=training, method="rf", preProcess="pca")    

}
```
```{r, echo=FALSE}
if (!file.exists('model.RData')) {
    stopCluster(cl)
    save(model, file='model.RData')
}
model
```

For this given model we can see very high accuracy error rate as low as 0.14%.

To do cross validation with random forest we can use `rfcv` function:
```{r, echo=FALSE}
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```
```{r}
rf.cv <- rfcv(subset(training, select=-classe) ,training$classe)
with(rf.cv,  plot(n.var, error.cv, log="x", type="o", lwd=2))
```
```{r, echo=FALSE}
stopCluster(cl)
```
